---
title: "Aim 1 Pilot Analyses"
author: "Ryan Gan"
date: "8/24/2017"
output: html_document
---

## Smoke as an effect modifier

This markdown file contains pilot data exploring the relationship between PM~2.5~ and cardiopulmonary hospital ED or urgent care visits. I am using Washington state CHARS data from 2012. I have time-stratified case-crossover designs. I include an interaction term between PM~2.5~ and a binary indicator of smoke.

Things I want to do: check distribution, try out a couple interaction models in case-crossover and time-series, assemble cohort to get some numbers.

```{r libraries}
library(tidyverse)
library(survival) 
```

Importing a couple of case-crossover datasets.

```{r import data, echo=F, message=F, warning=F}
#path <- "../.."
path <- "."
# pm2.5 estimates
pm <- read_csv(paste0(path,"/data/pm_data/zip_pm_to_merge_with_chars.csv")) %>% 
  na.omit() %>% 
  # make indicator variable
  mutate(gwr_smk0 = ifelse(geo_smk_pm> 0, 1, 0),
         gwr_smk5 = ifelse(geo_smk_pm> 5, 1, 0),
         gwr_smk10 = ifelse(geo_smk_pm > 10, 1, 0),
         gwr_smk15 = ifelse(geo_smk_pm > 15, 1, 0),
         gwr_smk20 = ifelse(geo_smk_pm > 20, 1, 0),
         gwr_10unit = geo_wt_pm/10)

# all cvd
cvd <- read_csv(paste0(path,"/data/health_data/",
                "cvd1_jul_to_oct_time_strat_casecross.csv")) %>% 
  # make indicator variable
  mutate(gwr_smk0 = ifelse(geo_smk_pm_zip > 0, 1, 0),
         gwr_smk5 = ifelse(geo_smk_pm_zip > 5, 1, 0),
         gwr_smk10 = ifelse(geo_smk_pm_zip > 10, 1, 0),
         gwr_smk15 = ifelse(geo_smk_pm_zip > 15, 1, 0),
         gwr_smk20 = ifelse(geo_smk_pm_zip > 20, 1, 0),
         gwr_10unit = geo_wt_pm_zip/10)
# heart failure
hf <- read_csv(paste0(path, "/data/health_data/",
                "hf1_jul_to_oct_time_strat_casecross.csv")) %>% 
  # make indicator variable
  mutate(gwr_smk0 = ifelse(geo_smk_pm_zip > 0, 1, 0),
         gwr_smk5 = ifelse(geo_smk_pm_zip > 5, 1, 0),
         gwr_smk10 = ifelse(geo_smk_pm_zip > 10, 1, 0),
         gwr_smk15 = ifelse(geo_smk_pm_zip > 15, 1, 0),
         gwr_smk20 = ifelse(geo_smk_pm_zip > 20, 1, 0),
         gwr_10unit = geo_wt_pm_zip/10)
# mi
mi <-  read_csv(paste0(path,"/data/health_data/",
                "mi1_jul_to_oct_time_strat_casecross.csv")) %>% 
  # make indicator variable
  mutate(gwr_smk0 = ifelse(geo_smk_pm_zip > 0, 1, 0),
         gwr_smk5 = ifelse(geo_smk_pm_zip > 5, 1, 0),
         gwr_smk10 = ifelse(geo_smk_pm_zip > 10, 1, 0),
         gwr_smk15 = ifelse(geo_smk_pm_zip > 15, 1, 0),
         gwr_smk20 = ifelse(geo_smk_pm_zip > 20, 1, 0),
         gwr_10unit = geo_wt_pm_zip/10)

# asthma
asthma <-  read_csv(paste0(path,"/data/health_data/",
                "asthma1_jul_to_oct_time_strat_casecross.csv")) %>% 
  # make indicator variable
  mutate(gwr_smk0 = ifelse(geo_smk_pm_zip > 0, 1, 0),
         gwr_smk5 = ifelse(geo_smk_pm_zip > 5, 1, 0),
         gwr_smk10 = ifelse(geo_smk_pm_zip > 10, 1, 0),
         gwr_smk15 = ifelse(geo_smk_pm_zip > 15, 1, 0),
         gwr_smk20 = ifelse(geo_smk_pm_zip > 20, 1, 0),
         gwr_10unit = geo_wt_pm_zip/10)
```

## PM~2.5~ characteristics

Background smoke for a given grid is the median value for that month when smoke is not over the grid, determined by HMS. I subtract backround smoke off the GWR estimate for the grid and population-weight by zipcode. In order to increase the specificity for PM~2.5~ due to wildfire smoke, I'll set a binary cutoff of >5 ug/m^3 above the GWR estimate minus background estimate to increase the specificity slightly.I will explore the best way to set a binary cutoff in the grant.

```{r pm distributions, echo=F, message=F, warning=F}
pm_mean <- pm %>% 
  group_by(gwr_smk5) %>% 
  summarise(pm_mean = mean(geo_wt_pm))

plot <- ggplot(data=pm, aes(x=geo_wt_pm, group = as.factor(gwr_smk5))) +
  geom_density(aes(fill = as.factor(gwr_smk5)), alpha = 0.5) +
  scale_fill_manual(values = c("red", "blue")) +
  geom_vline(data = pm_mean, aes(xintercept=pm_mean, group = gwr_smk5),
             color = c("red", "blue"), linetype = "dotted") +
  theme_bw()
# print plot
print(plot)
```

## Asthma

Exploring an interaction model between PM~2.5~ and the binary smoke indicator on asthma.

```{r asthma association}
mod <- clogit(outcome ~ gwr_10unit + gwr_smk5 +
          gwr_smk5*gwr_10unit  + wrf_temp_zip + 
            strata(PATIENTID), data =asthma)
n_obs <- mod$nevent
cov_matrix <- mod$var
# evidence for interaction
mod_tidy <- broom::tidy(mod)

# effect estimate of odds ratio table
asthma_table <- matrix(data = NA, nrow = 4, ncol = 5) %>% 
  as_tibble() %>% 
  rename(estimate = V1, 
         odds_ratio = V2,
         lower95 = V3,
         upper95 = V4,
         p_val = V5) %>% 
  mutate(outcome = "Asthma")

asthma_table[,1] <- c("pm_10unit", "smoke", "pm_diff","pm_smoke_10unit") 
# populate table 
asthma_table[1,2:4] <- round(exp(mod_tidy[1, c(2,6:7)]),2) # OR and 95CI
asthma_table[1,5] <- round(mod_tidy[1,5],2) # pval
asthma_table[2,2:4] <- round(exp(mod_tidy[2, c(2,6:7)]),2) # OR and 95CI
asthma_table[2,5] <- round(mod_tidy[2,5],2) # pval
# test for interaction
asthma_table[3,2:4] <- round(exp(mod_tidy[4, c(2,6:7)]),2) # OR and 95CI
asthma_table[3,5] <- round(mod_tidy[4,5],2) # pval

# linear combination estimate for smoke pm2.5
lin_combo <- c(1,0,0,1)
# estimate 
est <- as.numeric(t(lin_combo) %*% mod$coefficients)
# std error
se <- as.numeric(sqrt(t(lin_combo) %*% mod$var %*% lin_combo))
# calculate bounds
l95 <- est - (1.96*se)
u95 <- est + (1.96*se)
# calcuate z statistic and p val
z <- est/se
p <- 2*pnorm(-abs(z))
# fill in table
asthma_table[4, 2:4] <- round(exp(c(est, l95, u95)),2)
# p val
asthma_table[4,5] <- round(p,2)

knitr::kable(asthma_table, 
  caption = paste0("Asthma and PM2.5 effect modification estimates"))
```

The p-value for pm_diff (test for difference in CR function) may indicate marginal difference between effect estimates. The effect estimate for a 10 ug/m^3 increase in PM~2.5~ suggests a 10% increase for asthma, however the confidence interval suggests this is not a significant finding. In the context of other studies, this effect estimate is in line with the larger air pollution literature and I would conclude that it's not significant because it's power challenged. I expect that this term will be significant as we add addition years and states to the analysis.

Our indicator of a smoke day indicates that the risk for an asthma event on a smoke day is 26% greater than a non-smoke day and that this effect is significant.

The PM~2.5~ response on a smoke day indicates that a 10 ug/m^3 increase in PM~2.5~ increases risk for asthma by 7% and that this is a significant. This is the same magnitude of response we found in our Washington study when we subtracted off the baseline level.

```{r asthma power, eval = F}
n <- 1500
# assuming roughly 8% of smoke days
sim_data <- as_tibble(rbinom(n, 1, prob = 0.15)) %>% 
  rename(gwr_smk5 = value) %>%
  # simulate pm value without smoke and with smoke
  mutate(gwr_pm = ifelse(gwr_smk5 == 0, rnorm(n, mean = 5.7, sd = 2.3),
    rnorm(n, mean = 20.2, sd = 16.7)),
  # simulate logit response
  outcome = rbinom(n, 1, 1/(1+exp(-(-0.5 + 0.095*gwr_pm + 0.23*gwr_smk5 - 
            0.025*gwr_smk5*gwr_pm))))) 

sim_mod <- glm(outcome ~ gwr_pm + gwr_smk5 + gwr_pm*gwr_smk5, 
               family = "binomial", data = sim_data)

summary(sim_mod)
mod
summary(sim_data)

pm_sim_vals <- pm %>% 
  group_by(gwr_smk5) %>% 
  summarise(pm_mean = mean(geo_wt_pm), pm_sd = sd(geo_wt_pm))
pm_sim_vals
```

```{r std pwr}
# assuming base proportion of asthma is 0.06
0.06*.975-0.06
exp(-0.02523)

pr_exp <- prop.table(xtabs(~outcome + gwr_smk5, asthma),2)[2,]
pr_exp
# ratio
pr_exp[2]/pr_exp[1]  
# diff
pr_exp[2] - pr_exp[1]

test_mod <- clogit(outcome ~ gwr_smk5 +
            strata(PATIENTID), data =asthma)
test_mod
mod
# diff in probability
(1/exp(1+(-0.09))) - 1/exp(1+(0.025))
# vector of effect size
effect_size <- rep(seq(from=0.01,to=0.06, by=0.01),25)
# vector of sample sizes to estimate
case_size <- rep(seq(from = 1000, to = 25000, by = 1000) ,6)

power_df <- as_data_frame(cbind(effect_size, case_size))

  # apply function to calculate power
power_df$power <- apply(power_df, 1, function(x){
  pwr <- pwr::pwr.2p2n.test(h=x[1], n1=x[2], n2=x[2]*20, 
    sig.level=0.05, alternative = "two.sided")
  return(pwr$power)
  })

ggplot(data = power_df, aes(x = case_size, y = power, 
  colour = as.factor(effect_size))) +
  geom_line() +
  guides(colour=guide_legend("Effect Size")) +
  ylab("Power") +
  xlab("Case Size") +
  theme_bw()

```

## All CVD

```{r cvd association}
mod <- clogit(outcome ~ gwr_10unit + gwr_smk5 +
          gwr_smk5*gwr_10unit  + wrf_temp_zip + 
            strata(PATIENTID), data =cvd)

# evidence for interaction
mod_tidy <- broom::tidy(mod)
mod
mod$nevent
# cvd events on a smoke day
cvd_smoke <- cvd %>% 
  group_by(gwr_smk5) %>% 
  summarise(n_case = sum(outcome))
cvd_smoke
# effect estimate of odds ratio table
cvd_table <- matrix(data = NA, nrow = 4, ncol = 5) %>% 
  as_tibble() %>% 
  rename(estimate = V1, 
         odds_ratio = V2,
         lower95 = V3,
         upper95 = V4,
         p_val = V5) %>% 
  mutate(outcome = "CVD")

cvd_table[,1] <- c("pm_10unit", "smoke", "pm_diff","pm_smoke_10unit") 
# populate table 
cvd_table[1,2:4] <- round(exp(mod_tidy[1, c(2,6:7)]),2) # OR and 95CI
cvd_table[1,5] <- round(mod_tidy[1,5],2) # pval
cvd_table[2,2:4] <- round(exp(mod_tidy[2, c(2,6:7)]),2) # OR and 95CI
cvd_table[2,5] <- round(mod_tidy[2,5],2) # pval
# test for interaction
cvd_table[3,2:4] <- round(exp(mod_tidy[4, c(2,6:7)]),2) # OR and 95CI
cvd_table[3,5] <- round(mod_tidy[4,5],2) # pval

# linear combination estimate for smoke pm2.5
lin_combo <- c(1,0,0,1)
# estimate 
est <- as.numeric(t(lin_combo) %*% mod$coefficients)
# std error
se <- as.numeric(sqrt(t(lin_combo) %*% mod$var %*% lin_combo))
# calculate bounds
l95 <- est - (1.96*se)
u95 <- est + (1.96*se)
# calcuate z statistic and p val
z <- est/se
p <- 2*pnorm(-abs(z))
# fill in table
cvd_table[4, 2:4] <- round(exp(c(est, l95, u95)),2)
# p val
cvd_table[4,5] <- round(p,2)

knitr::kable(cvd_table, 
  caption = paste0("CVD and PM2.5 effect modification estimates"))
```

For CVD hospitalizations, there is no association with PM~2.5~ and the magnitude is in the opposite direction that what we'd expect. Our smoke term indicates a lower likelihood for risk of CVD hospitalization on a smoke day (may be people saying indoors). The interaction term (pm_dff) may suggest a difference in effect. However, when we look at the PM~2.5~ term on a smoke day, there is no association.

## Heart Failure

```{r hf association}
mod <- clogit(outcome ~ gwr_10unit + gwr_smk5 +
          gwr_smk5*gwr_10unit  + wrf_temp_zip + 
            strata(PATIENTID), data = hf)

# evidence for interaction
mod_tidy <- broom::tidy(mod)
mod_tidy
# effect estimate of odds ratio table
hf_table <- matrix(data = NA, nrow = 4, ncol = 5) %>% 
  as_tibble() %>% 
  rename(estimate = V1, 
         odds_ratio = V2,
         lower95 = V3,
         upper95 = V4,
         p_val = V5) %>% 
  mutate(outcome = "Heart Failure")

hf_table[,1] <- c("pm_10unit", "smoke", "pm_diff","pm_smoke_10unit") 
# populate table 
hf_table[1,2:4] <- round(exp(mod_tidy[1, c(2,6:7)]),2) # OR and 95CI
hf_table[1,5] <- round(mod_tidy[1,5],2) # pval
hf_table[2,2:4] <- round(exp(mod_tidy[2, c(2,6:7)]),2) # OR and 95CI
hf_table[2,5] <- round(mod_tidy[2,5],2) # pval
# test for interaction
hf_table[3,2:4] <- round(exp(mod_tidy[4, c(2,6:7)]),2) # OR and 95CI
hf_table[3,5] <- round(mod_tidy[4,5],2) # pval

# linear combination estimate for smoke pm2.5
lin_combo <- c(1,0,0,1)
# estimate 
est <- as.numeric(t(lin_combo) %*% mod$coefficients)
# std error
se <- as.numeric(sqrt(t(lin_combo) %*% mod$var %*% lin_combo))
# calculate bounds
l95 <- est - (1.96*se)
u95 <- est + (1.96*se)
# calcuate z statistic and p val
z <- est/se
p <- 2*pnorm(-abs(z))
# fill in table
hf_table[4, 2:4] <- round(exp(c(est, l95, u95)),2)
# p val
hf_table[4,5] <- round(p,2)

knitr::kable(hf_table, 
  caption = paste0("Heart failure and PM2.5 effect modification estimates"))
```

For heart failure, we do see an increase in risk for hospitalization with a 10 ug/m^3 increase and on smoke days. The test for difference may suggest a difference in effect, where increases in PM~2.5~ on a smoke day is not associated with heart failure. Some interpretations of this may be that the initial risk is just the smoke itself, with increases in concentrations at this point having no effect. This would be probably the most interesting interaction to evaluate in the large study.

## Myocardial Infarction

```{r mi association}
mod <- clogit(outcome ~ gwr_10unit + gwr_smk5 +
          gwr_smk5*gwr_10unit  + wrf_temp_zip + 
            strata(PATIENTID), data = mi)

# evidence for interaction
mod_tidy <- broom::tidy(mod)
mod_tidy
# effect estimate of odds ratio table
mi_table <- matrix(data = NA, nrow = 4, ncol = 5) %>% 
  as_tibble() %>% 
  rename(estimate = V1, 
         odds_ratio = V2,
         lower95 = V3,
         upper95 = V4,
         p_val = V5) %>% 
  mutate(outcome = "Myocardial Infarction")

mi_table[,1] <- c("pm_10unit", "smoke", "pm_diff","pm_smoke_10unit") 
# populate table 
mi_table[1,2:4] <- round(exp(mod_tidy[1, c(2,6:7)]),2) # OR and 95CI
mi_table[1,5] <- round(mod_tidy[1,5],2) # pval
mi_table[2,2:4] <- round(exp(mod_tidy[2, c(2,6:7)]),2) # OR and 95CI
mi_table[2,5] <- round(mod_tidy[2,5],2) # pval
# test for interaction
mi_table[3,2:4] <- round(exp(mod_tidy[4, c(2,6:7)]),2) # OR and 95CI
mi_table[3,5] <- round(mod_tidy[4,5],2) # pval

# linear combination estimate for smoke pm2.5
lin_combo <- c(1,0,0,1)
# estimate 
est <- as.numeric(t(lin_combo) %*% mod$coefficients)
# std error
se <- as.numeric(sqrt(t(lin_combo) %*% mod$var %*% lin_combo))
# calculate bounds
l95 <- est - (1.96*se)
u95 <- est + (1.96*se)
# calcuate z statistic and p val
z <- est/se
p <- 2*pnorm(-abs(z))
# fill in table
mi_table[4, 2:4] <- round(exp(c(est, l95, u95)),2)
# p val
mi_table[4,5] <- round(p,2)

knitr::kable(mi_table, 
  caption = paste0("Myocardial infarction and PM2.5 effect modification estimates"))
```

No evidence of association for either main term, but difference may indicate a difference in effect for PM~2.5~ on a smoke day. OR suggests that a 10 ug/m^3 increase on a smoke day is associated with a 3% increase in MI hospitalization, although not significant. Could very well be a power issue.

Figure of pilot data for grant.

```{r pilot plot}
small_multiples_df <- rbind(asthma_table, cvd_table, hf_table, mi_table)

interaction_p <- small_multiples_df %>%  filter(estimate == "pm_diff") %>% 
  select(outcome, p_val) %>% 
  rename(p_int = p_val)

or_df <- small_multiples_df %>% 
  filter(estimate == c("pm_10unit", "pm_smoke_10unit")) %>% 
  right_join(interaction_p, by = "outcome") %>% 
  mutate(outcome_p = paste0(outcome, "\n Interaction p-value = ",p_int))

or_plot <- ggplot(data = or_df, 
  aes(x = estimate, y = odds_ratio, colour = estimate)) +
  scale_color_manual(name = expression(paste("PM"[2.5], " Group")),
                     values = c("blue", "red"),
                     labels = c("Anthropogenic", "Wildfire")) +
  geom_point() +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2) +
  ylab(expression(paste("Odds Ratio for 10 µg/m"^3, " Increase in PM"[2.5]))) +
  facet_wrap(~outcome_p, scales = "free") +
   # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    #strip.text = element_text(size = 8),
    # axis element
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

or_plot
```
