---
title: "Distributed Lag Model"
author: "Ryan_Gan"
date: "January 24, 2017"
output: html_document
---


## Purpose of Markdown File

Trying out distributed lag models using the 'dlnm' package. Loading libraries.

```{r libraries used, echo = F, message = F}

library(tidyverse) # data wrangle package
library(dlnm) # distributed lag models
library(lme4) # random-effects model
library(broom) # broom for tidy data from stats models
library(splines) # spline library
library(rmeta) # meta analysis
library(mvmeta) # multivariate meta analysis for pooling region-specific dlnm
```

Importing county-level time-series data.

```{r county time series import, echo = F, message = F}
# file path
file_path <- paste0("../../data/health_data/wa_2012_county_time_series.csv")

# read file
wash_ts_df <- read_csv(file_path)

summary(wash_ts_df)

# read in county estimates for offset
pop_path <- paste0("./data/health_data/wa_census_pop_est_county.csv")
pop_est <- read_csv(pop_path) %>% 
  # seperate county from state
  separate(GEO.display.label, c("county", "state"), sep = ",") %>% 
  # remove the word 'county' from the county string (mind the space)
  mutate(county = gsub(county, pattern = " County", replacement = "")) %>% 
  # rename variables
  rename(fips = GEO.id2, pop_2010 = respop72010, pop_2011 = respop72011,
         pop_2012 = respop72012, pop_2013 = respop72013, pop_2014 = respop72014,
         pop_2015 = respop72015, pop_2016 = respop72016) %>% 
  # limit to certain variables
  select(state, county, fips, pop_2010:pop_2016)

# only need 2012 population estimate as an offset
county_pop_2012 <- pop_est %>% select(county, pop_2012)

# analysis dataframe
wash_jul_oct_ts_df <- wash_ts_df %>% 
  # restricting points
  filter(date >= "2012-07-01" & date <= "2012-10-31") %>% 
  # exclude 'Unknown' and 'not_wa_residence'
  filter(county != "not_wa_residence" & county != "Unknown") %>% 
  # create day of the week variable
  mutate(day = as.factor(weekdays(date))) %>% 
  # set missing outcome values to 0 since missing indicates no ER or urgent care 
  # visits on that date (which is reasonable in sparsely populated counties)
  mutate_each(funs(wo_miss = ifelse(is.na(.), 0, .)), n_obs:ra_n) %>% 
  # create a binary smoke indicator variable based smk variables >5, >10, and >15 units 
  mutate(wrf_smk5 = ifelse(wrf_smk_pm >= 5, 1, 0), 
         wrf_smk10 = ifelse(wrf_smk_pm >= 10, 1, 0),
         wrf_smk15 = ifelse(wrf_smk_pm >= 15, 1, 0), 
         geo_smk5 = ifelse(geo_smk_pm >= 5, 1, 0),
         geo_smk10 = ifelse(geo_smk_pm >= 10, 1, 0),
         geo_smk15 = ifelse(geo_smk_pm >= 15, 1, 0),
         krig_smk5 = ifelse(krig_smk_pm >= 5, 1, 0),
         krig_smk10 = ifelse(krig_smk_pm >= 10, 1, 0),
         krig_smk15 = ifelse(krig_smk_pm >= 15, 1, 0), 
         season = ifelse(date >= "2012-06-22" &  date <= "2012-09-22", "summer",
                  ifelse(date >= "2012-09-23" & date <= "2012-12-21", "fall",
                          "other"))) %>% 
  # rescale pm variables to 10 units
  mutate_each(funs(unit10 = ./10), wrf_pm:krig_smk_pm) %>% 
  select(1:2, 31:33, 36, 34:35, 38:44, 28:29, 30, 55, 58, 
         65, 64, 63) %>% # subset and order variables used in analysis
  # join county population estimates for offset
  right_join(county_pop_2012, by = "county")

```

## Visualizing Time-Series Data

First step is to visualize how both the health outcomes and our smoke concentrations vary over time.

### Health Outcomes for the Full Year of 2012

First plot contains a small-multiples plot of health outcomes of interest that aggregates all county-specific time-series for the entire state. There is no apparent pattern over the July 1st to October 31st timeseries for CVD outcomes. It's generally hard to observe these patterns in some of the sparsley populationed counties where most smoke occured, so I like to aggregate for the state to get an idea of the background counts.

```{r plot of events over time, echo = F, message = F, results="asis"}

# wrangle time-series dataframe to create small multiples of health outcomes
outcome_plot_df <- wash_ts_df %>% filter(county != "Unknown") %>% # remove unknown county
  select(1:8, 10:16) # select columns of health outcomes

wa_state_aggregate <- outcome_plot_df %>% select(-county) %>% 
  group_by(date) %>% 
  summarise_each(funs(sum(., na.rm = T))) %>% 
  gather(key = outcome, value = n, -date)

#summary(wa_state_aggregate)

ggplot(wa_state_aggregate, aes(x = date, y = n)) +
  geom_point() +
  facet_wrap(~outcome, scales = "free")
  
```

### Health Outcome Time Series from July 1st to October 31st 2012

Plot of the time trends in outcomes for just the study period of interest.

```{r plot of events jul to oct, echo = F, message = F, results="asis"}

# wrangle time-series dataframe to create small multiples of health outcomes
outcome_plot_df <- wash_ts_df %>% filter(county != "Unknown") %>% # remove unknown county
  select(1:8, 10:16) # select columns of health outcomes

wa_state_aggregate_jul_oct <- outcome_plot_df %>% select(-county) %>% 
  group_by(date) %>% 
  summarise_each(funs(sum(., na.rm = T))) %>% 
  gather(key = outcome, value = n, -date) %>% 
  filter(date >= "2012-07-01" & date <= "2012-10-31")

#summary(wa_state_aggregate)

ggplot(wa_state_aggregate_jul_oct, aes(x = date, y = n)) +
  geom_point() +
  facet_wrap(~outcome, scales = "free")

```  

It appears that there may be some temporal trends with some of the respiratory conditions. Asthma is a good example, where the number in the state is around 10 in July/August, but creeps up to as high as 30 in mid-September and early-October. Could part of that increase be due to wildfire smoke or is it part of a natural trend?


### County Smoke Concentrations and Temperature Over Time

Most counties in the state are not impacted by smoke. However, some of the central counties do have the expected peak of high smoke concentrations around mid September.

```{r exposure series, echo = F, message = F, results="asis"}
# check time series of smoke in each county
ggplot(wash_jul_oct_ts_df, aes(x = date, y = geo_smk_pm_unit10)) +
  geom_point() +
  facet_wrap(~county)

# temp
ggplot(wash_jul_oct_ts_df, aes(x = date, y = wrf_temp)) +
  geom_point() +
  facet_wrap(~county)
```

### Asthma and Smoke Time-Series Analysis

Counts of asthma claims tend to tick up a bit in the Fall, right around the state was affected by smoke. First step is to account for the temporal trends in asthma.

Plotting the asthma counts for the overall state and then each county for the July 1st to October 31st 2012 time series.

```{r study period asthma counts by county, message=F, echo=F, results="asis"}
# aggregated counts limited to July 1 to Oct 31st 2012
asthma_state_aggregate_jul_oct <- wa_state_aggregate %>% 
  filter(date >= "2012-07-01" & date <= "2012-10-31") %>% 
  filter(outcome == "asthma_n")

ggplot(asthma_state_aggregate_jul_oct, aes(x = date, y = n)) +
  geom_point() +
  ggtitle("Washington asthma counts for the 2012 fire season")

# counties small multiple plot of asthma count
ggplot(wash_jul_oct_ts_df, aes(x = date, y = asthma_n_wo_miss)) +
  geom_point() +
  facet_wrap(~county) +
  ggtitle("Asthma counts for the 2012 fire season by county")

```

It looks like in some of the bigger counties, asthma counts begin to increase in mid August. In King county (where Seattle is), the trend may even suggest that the peak is mid September (around the same time as the fire), and then beings to decrease. King county wasn't really affected by smoke, so it would be safe to say this is shape likely represents the background trend in asthma. It is hard or impossible to see trends in the more sparsely populated counties. Maybe this would be a good place to try and use empirical Bayes estimation? I should ask Ander.

There is evidence of overdispersion in each county. General Poisson regression may not be appropriate. 

Another issue is how to deal with any potential variation in asthma counts by county. Off the top of my head, it can either be ajusted for as a fixed effect, treated as a random effect, run each counties as seperate models and pool results in a meta-analysis. 


### GLM Poisson without Natrual Spline

Exploring very basic models.

```{r basic glm regression, echo = F, message = F, results="asis"}
summary(wash_jul_oct_ts_df)
# crude model with Poisson distribution
pois_mod <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10 + 
                  offset(log(pop_2012)), wash_jul_oct_ts_df, family="poisson")
# print mod
tidy(pois_mod)

# dispersion parameter for a poisson regression is assumed to be 1
summary(pois_mod)$dispersion

# crud model with Quasi-Poisson distribution for overdispersion
quasipois_mod <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10 +
                     offset(log(pop_2012)), wash_jul_oct_ts_df, 
            family="quasipoisson")
# print mod
tidy(quasipois_mod)
# dispersion parameter for quasipoisson
summary(quasipois_mod)$dispersion

# dispersion parameter for quasipoisson is >1, suggesting overdispersion

# beta estimate is not different, but standard error is higher in quasipoisson
```

Based on just these results, it's best to use the quasipoisson distribution to account for zero-inflated values. Beta estimate doesn't change much, but standard error is different enough where it could lead to different conclusions of just Poisson was used.

All subsequent models will use quasipoisson family. Next model adjusts for temperature and county

```{r glm adjusting for temperature, echo = F, message = F, results="asis"}

# adjusted for temp
quasipois_mod <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10 + wrf_temp +
                     offset(log(pop_2012)), 
                     wash_jul_oct_ts_df, family="quasipoisson")
# print mod
tidy(quasipois_mod)
# dispersion parameter
summary(quasipois_mod)$dispersion
# still overdispersed adjusting for temp

# adjusting for county as a fixed effect
quasipois_mod <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10 + wrf_temp + 
                     county + offset(log(pop_2012)), wash_jul_oct_ts_df, 
                     family="quasipoisson")
# print mod
tidy(quasipois_mod) 
# dispersion parameter
summary(quasipois_mod)$dispersion
# when accounting for county, dispersion is actually pretty close to 1
# suggests it may be okay to use a standard poisson distributions as
# long as county is accounted for

```

Summary from above models is that once county is accounted for, a 10 ug/m^3 increase in smoke PM2.5 is associated with a ~11% increase in asthma hospitalizations for ED or urgent care. Also, when county is accounted for, the dispersion parameter is approximaly 1. This suggests that it may be okay to use a Poisson distribution.

I'm going to try a random effect model with a Poisson distribution, as quasipoisson family does not exist for mixed model package.

```{r time series with a mixed model, echo = F, message = F, results="asis"}
# convergence issues when temp is in the model. going to standardize
wash_jul_oct_ts_df$wrf_temp_sd <- scale(wash_jul_oct_ts_df$wrf_temp,
                                        center = T, scale = T)

# standarizing fixes the warning messages of the generalized mixed mod

# asthma model where county is a random effect instead of fixed
mixed_mod <- glmer(asthma_n_wo_miss ~ geo_smk_pm_unit10 + wrf_temp_sd + 
                   offset(log(pop_2012)) + (1|county), wash_jul_oct_ts_df, 
                   family = "poisson")
# full output
summary(mixed_mod)

# reduced output
tidy(mixed_mod)

```

These results are almost identical to the model where county is treated as a fixed effect. Increases in a 10 ug/m^3 of wildfire smoke PM2.5 is associated with an ~11% increase in asthma hospitalizations.

### Pooled County-Specific Time Series 

This approch runs a time series for each county, then pools the results in a meta-analysis. I expect this will yield very similar results to the above results.


```{r meta of county time series, echo = F, message = F, results="asis"}

county_mods <- wash_jul_oct_ts_df %>% 
  group_by(county) %>% # group by county
  do(geo_smk_beta = glm(asthma_n_wo_miss ~ geo_smk_pm_unit10 + wrf_temp_sd +
                        offset(log(pop_2012)), data = ., family="quasipoisson")) 
  

county_betas <- tidy(county_mods, geo_smk_beta) %>% 
  filter(term == "geo_smk_pm_unit10")

# meta analysis results for asthma and smoke
meta_results <- meta.summaries(county_betas$estimate, county_betas$std.error, 
                               method = "fixed")

summary(meta_results)
meta_results

```

This meta-analysis approach yields very similar results where there is an 11% increase in asthma hospitalizations associated with a 10 ug/m^3 increase in wildfire smoke PM2.5.

### GLM Poisson with Natrual Spline

I will now fit a natural spline using the 'splines' package to see if it improves the model. There are a couple ways to do it, but 

```{r asthma association with splines, echo = F, message = F, results="asis"}
# aggregate outcomes for the entire state and average/median smoke estimates
outcomes_df <- wash_jul_oct_ts_df %>% 
  select(date, contains("wo_miss")) %>% 
  group_by(date) %>% 
  summarise_each(funs(sum(.)))

pm_df <- wash_jul_oct_ts_df %>% 
  select(date, wrf_temp, contains("pm")) %>% 
  group_by(date) %>% 
  summarise_each(funs(mean, median))

wash_outcome_counts_pm <- outcomes_df %>% 
  full_join(pm_df, by = "date") %>% 
  mutate(day = weekdays(date),
    season = ifelse(date >= "2012-06-22" &  date <= "2012-09-22", "summer",
      ifelse(date >= "2012-09-23" & date <= "2012-12-21", "fall", "other")))

summary(wash_outcome_counts_pm)

# define the function of the spline for the asthma trend for the entrie state
asthma_spl <- ns(wash_outcome_counts_pm$date, df = 3)
# general question. How does the spline know the trend is for asthma and not 
# other outcomes? very curious... I will need to look in to this

# model n asthma events with a spline with 7 df (knotches)
asthma_mod <- glm(asthma_n_wo_miss ~ asthma_spl,
                  wash_outcome_counts_pm, family="poisson") 

# check spline fit to asthma
summary(asthma_mod)

# join predicted number of events to a new df for plotting
asthma_pred_df <- wash_outcome_counts_pm %>% 
  select(date, asthma_n_wo_miss) %>% 
  mutate(expected_n = predict(asthma_mod, type="response"),
         res = residuals(asthma_mod, type="response"))

# plot predicted events over outcomes
ggplot(asthma_pred_df, aes(x = date, y = asthma_n_wo_miss)) +
  geom_point() +
  stat_smooth(aes(y = expected_n), se = F, method = "loess") +
  theme_bw()

# plot residuals
ggplot(asthma_pred_df, aes(x = date, y = res)) +
  geom_point() +
  geom_hline(yintercept = 0, colour = "blue") +
  theme_bw()


# run the regression models with the spline
# models for association between smoke pm and daily asthma count 
# unadjusted model
asthma_mod_unadj <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10_median + 
                        wrf_temp_mean + day + season, 
                        wash_outcome_counts_pm, family="quasipoisson")
summary(asthma_mod_unadj)

# controlling for seasonality with spline
asthma_mod_adj <- update(asthma_mod_unadj,.~.+ asthma_spl)
summary(asthma_mod_adj)

```

When using a spline, there is no association between increasing smoke exposure and asthma counts. However, I had to aggregate to the state. One solution is to fit a spline to each county and run county-specific models and then aggregate using meta-analysis.


## Distributed Lag Models

General interpretation would be for a 1 ug/m^3 increase in smoke PM2.5, asthma events increase by 1%.

```{r simple dlm, echo = F, message = F, results="asis"}
cb1.pm <- crossbasis(wash_ts_df$geo_wt_pm, lag = 15, argvars=list(fun="lin"),
                     arglag=list(fun="poly", degree=3))

summary(cb1.pm)


model1 <- glm(asthma_n ~ cb1.pm +wrf_temp,
              family=quasipoisson(), wash_ts_df)

summary(model1)


pred1.pm <- crosspred(cb1.pm, model1, at=0:20, bylag=0.2, cumul=TRUE)

plot(pred1.pm, "slices", var=10, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
main="Association with a 1-unit increase in PM2.5")


plot(pred1.pm, "slices", var=10, col=2, cumul=TRUE, ylab="Cumulative RR",
main="Cumulative association with a 1-unit increase in PM2.5")


```

May not be appropriate to aggregate for the entrie state. Just based on outcome and smoke plots, it's clear that there is heterogenity by county.

### Two-Stage Distributed Lag by County and Meta-Analysis

Following two-stage approach where a DLM is run for each county, then pooled via meta-analysis.

```{r two stage prep, echo = F, message = F, results="asis"}
# create county vector
counties <- as.character(unique(wash_jul_oct_ts_df$county))

# create dataframe list of each county
county_list_df <- lapply(counties, function(x) 
  wash_jul_oct_ts_df[wash_jul_oct_ts_df$county==x,])
# assign each element of the list object the county name
names(county_list_df) <- counties
# calculate length of counties (39)
m <- length(counties)

# GWR PM2.5 ranges
smk_range <- t(sapply(county_list_df, 
  function(x) range(x$geo_smk_pm_unit10,na.rm=T)))

# -----

# function to compute the QAIC (quasi-poisson AIC)
fqaic <- function(model) {
  loglik <- sum(dpois(model$y,model$fitted.values,log=TRUE))
  phi <- summary(model)$dispersion
  qaic <- -2*loglik + 2*summary(model)$df[3]*phi
  return(qaic)
}

```

I'm going to run a single county model for Chelan county.

```{r chelan dlm linear, echo = F, message = F, results="asis"}
# select county
county_name <- "Chelan"

# arguments and lists for cross-basis
bound <- smk_range[4,]
varknots <- bound[1] + diff(bound)/3*(1:2)
argvar <- list(fun="lin")
arglag <- list(fun="ns",df=4)

# cross-basis for smoke
# - linear predictor between smoke and outcome
# - natural cubic spline for lag, with DF at equally-spaced log vals
# warning for prediction beyond boundaries suppressed 
suppressWarnings(
cb_chelan <- crossbasis(county_list_df[[county_name]]$geo_smk_pm_unit10,
                 lag=10, argvar=argvar, arglag=arglag)
)
summary(cb_chelan)


# run the model
model <- glm(asthma_n_wo_miss ~ cb_chelan + ns(date, df=3) + wrf_temp,
  family = "quasipoisson", county_list_df[[county_name]])

summary(model)

# prediction using:
#   crosspred for bi-dimensional reslationship 
#   crossreduce for uni-dimensional summaries 
# (NB: all the estimates are presented by crosspred alone)

pred_smk <- crosspred(cb_chelan, model, at = 0:20, bylag = 0.2, cumul = T)
pred_smk$matRRfit

# plot the lag-response curves for specific and incremental cumulative effects
plot(pred_smk, "slices", var=10, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
main="Lag-response curve for a 10-unit increase in Smoke PM10")


plot(pred_smk, "slices", var=10, col=2, cumul=TRUE, ylab="Cumulative RR",
main="Lag-response curve of incremental cumulative effects")


```


```{r first stage county prediction, echo = F, message = F, results="asis"}
#  these first two commands aren't necissary since I'm assuming linear for now

# assuming linear relationship between smoke and outcomes
argvar <- list(fun="lin")
# fitting a natural spline for lag with 5 knots?
arglag <- list(fun="ns", df=4)
lag <- 10

# ----
# build matrix where results will be stored
#   y- matrix for outcome parameters
#   S- list of (co)variance matrices

# overall cumulative summaries
# need 4 columns for the 4 degrees of freedom for the spline
asthma_all <- matrix(NA, length(county_list_df), 4, 
               dimnames=list(counties,paste("b",seq(4),sep="")))

# (CO)VARIANCE MATRICES
asthma_cov_all <- vector("list",length(county_list_df))
names(asthma_cov_all) <- counties

# Q-AIC
qaic <- 0

# loop for each city ----

# warning for predictions out of bounds suppressed
system.time({
for(i in seq(county_list_df)) {

  # LOAD
  sub <- county_list_df[[i]]
  lag
  argvar
  arglag
  # DEFINE THE CROSS-BASES
  suppressWarnings({
  asthma_cb <- crossbasis(sub$geo_smk_pm_unit10,
                     lag=lag, argvar=argvar, arglag=arglag)

  })

  # RUN THE FIRST-STAGE MODELS
  model <- glm(asthma_n_wo_miss ~ asthma_cb + offset(log(pop_2012)), #+ ns(date, df=3),
    family = "quasipoisson", sub)
  
  summary(model)
  
# ----
  # REDUCTION TO SUMMARY ASSOCIATIONS

  # prediction summary for smoke 10 ug/m^3
  suppressWarnings({
    asthma_cr_all <- crossreduce(asthma_cb, model, type = "var", value = 1)

  })

#----
  # STORE THE RESULTS

  # OVERALL CUMULATIVE SUMMARY FOR THE MAIN MODEL
  asthma_all[i,] <- coef(asthma_cr_all)
  asthma_cov_all[[i]] <- vcov(asthma_cr_all)

  # Q-AIC
  qaic[i] <- fqaic(model)
}
})

# GRAND Q-AIC
sum(qaic) 

#

```

Asthma outcome.

```{r dlm in a mixed model, echo = F, message = F, results="asis"}


####################################################################
# PERFORM MULTIVARIATE META-ANALYSIS

# SELECT THE ESTIMATION METHOD
method <- "reml"
# IN THE CURRENT VERSION, SET control=list(trace=6,REPORT=1) TO 
#   INSPECT THE OPTIMIZATION SEARCH

# OVERALL CUMULATIVE SUMMARY FOR THE MAIN MODEL
mv_asthma_all <- mvmeta(asthma_all~1,asthma_cov_all,method=method)
summary(mv_asthma_all)

####################################################################
# CREATE BASES FOR PREDICTION
asthma_cr_all$RRlow
# BASES OF TEMPERATURE AND LAG USED TO PREDICT, EQUAL TO THAT USED FOR ESTIMATION
# COMPUTED USING THE ATTRIBUTES OF THE CROSS-BASIS USED IN ESTIMATION
xlag <- 0:110/10
blag <- do.call("onebasis",c(list(x=xlag),attr(asthma_cb,"arglag")))

####################################################################
# REGION-SPECIFIC FIRST-STAGE SUMMARIES

reg_county_all <- apply(asthma_all,1 , function(x) exp(blag%*%x))


####################################################################
# PREDICTION FOR A GRID OF TEMPERATURE AND LAG VALUES

# PREDICTOR-SPECIFIC SUMMARIES FOR 10 ug/m^3 of smoke
cp_smk10 <- crosspred(blag, coef=coef(mv_asthma_all),
                      vcov=vcov(mv_asthma_all),
                      model.link="log", at = 0:110/10)

# summary stats
cp_smk10$allRRfit
cp_smk10$allRRlow
cp_smk10$matRRfit

# plot
plot.crosspred(cp_smk10, type="n", ylim=c(.5, 2.0), ylab="RR", xlab="Lag")
matplot(xlag, reg_county_all, type="l", col=grey(0.5),lty=2,add=T)
abline(h=1)
lines(cp_smk10,col=2,lwd=2)
legend ("top",c("Pooled (with 95%CI)","First-stage region-specific"),
  lty=c(1,2),lwd=1.5,col=c(2,grey(0.7)),bty="n",inset=0.1,cex=0.8)
mtext(text=paste("Predictor-specific summary"),cex=1)

# this seems to work.
```

I think I figured out the DLM without non-linear relationships. I'm going to run it for all respiratory and all CVD outcomes for the abstract.

### Two-Stage DLM for All Respiratory and CVD Outcomes

```{r resp and cvd two stage analysis }
# setup ----
# assuming linear relationship between smoke and outcomes
argvar <- list(fun="lin")
# fitting a natural spline for lag with 5 knots?
arglag <- list(fun="ns", df=4)
lag <- 10

# build matrix where results will be stored
#   resp_all- matrix for outcome parameters
#   resp_cov_all - list of (co)variance matrices

# overall cumulative summaries
# need 4 columns for the 4 degrees of freedom for the spline
cvd_all <- resp_all <- matrix(NA, length(county_list_df), 4, 
  dimnames=list(counties,paste("b",seq(4),sep="")))

# (co)variance matrices
cvd_cov_all <- resp_cov_all <- vector("list",length(county_list_df))
names(resp_cov_all) <- counties
names(cvd_cov_all) <- counties

# Q-AIC
qaic_resp <- 0
qaic_cvd <- 0

# loop for each city ----

# warning for predictions out of bounds suppressed
system.time({
for(i in seq(county_list_df)) {

  # load county dataframe
  sub <- county_list_df[[i]]

  # define the cross-basis for smoke ----
  suppressWarnings({
  smk_cb <- crossbasis(sub$geo_smk_pm_unit10,
                     lag=lag, argvar=argvar, arglag=arglag)
  
  temp_cb <- crossbasis(sub$wrf_temp, lag = lag, argvar = argvar,
                        arglag = arglag)
  
  })


  # run the first stage models
  resp_model <- glm(resp_n_wo_miss ~ smk_cb + temp_cb + 
                    offset(log(pop_2012)), 
    family = "quasipoisson", sub)
  
  cvd_model <- glm(cvd_n_wo_miss ~ smk_cb + temp_cb + 
                  offset(log(pop_2012)), 
    family = "quasipoisson", sub)
  
  # reduction to summary results ---
  # prediction summary for smoke 10 ug/m^3 
  suppressWarnings({
    resp_cr_all <- crossreduce(smk_cb, resp_model, type = "var", value = 1)
    
    cvd_cr_all <- crossreduce(smk_cb, cvd_model, type = "var", value = 1)

  })


  # store results in matrix ---

  # daily lag results
  resp_all[i,] <- coef(resp_cr_all)
  resp_cov_all[[i]] <- vcov(resp_cr_all)
  
  cvd_all[i,] <- coef(cvd_cr_all)
  cvd_cov_all[[i]] <- vcov(cvd_cr_all)

  # Q-AIC
  qaic_resp[i] <- fqaic(resp_model)
  qaic_cvd[i] <- fqaic(cvd_model)
}
})

# GRAND Q-AIC
sum(qaic_resp)
sum(qaic_cvd)

# garfield county producing error for cvd; fails to converge

# second stage ----
# multivariate meta analysis

# reml estimation method
method <- "reml"

# overall pooled summary for resp and cvd model
mv_resp_all <- mvmeta(resp_all~1, resp_cov_all, method = method)

# taking out garfiled county as it will affect the convergence of
# the meta analysis 
cvd_wo_garfield <- subset(cvd_all, rownames(cvd_all) != "Garfield")
cvd_cov_all <- cvd_cov_all

#counties_wo_garfield <- counties[c(1:11, 13:39)]
#counties_wo_garfield

cvd_cov_wo_garfield <- cvd_cov_all[c(1:11, 13:39)]

mv_cvd_all <- mvmeta(cvd_wo_garfield~1, 
                     cvd_cov_wo_garfield, method = method)

# # meta summaries
# summary(mv_resp_all)
# summary(mv_cvd_all)

# create bases for prediction ----

# bases of smoke lag
xlag <- 0:110/10
blag <- do.call("onebasis",c(list(x=xlag),attr(smk_cb,"arglag")))

# county-specific first stage summaries

resp_county_all <- apply(resp_all,1, function(x) exp(blag%*%x))
cvd_count_all <- apply(cvd_wo_garfield, 1, function(x) exp(blag%*%x))


# prediction for a grid of lag days

# predictor-specific summaries for 10 ug/m^3 of smoke
resp_cp_smk10 <- crosspred(blag, coef=coef(mv_resp_all),
                      vcov=vcov(mv_resp_all),
                      model.link="log", at = 0:110/10)


cvd_cp_smk10 <- crosspred(blag, coef=coef(mv_cvd_all), 
                          vcov=vcov(mv_cvd_all),
                          model.link="log", at = 0:110/10)

# # summary stats for resp
resp_cp_smk10$allRRfit
resp_cp_smk10$allRRlow
resp_cp_smk10$allRRhigh
# # summary stats for cvd
cvd_cp_smk10$allRRfit
cvd_cp_smk10$allRRlow
cvd_cp_smk10$allRRhigh

# plot of all resp
plot.crosspred(resp_cp_smk10, type="n", ylim=c(.8, 1.2), ylab="RR", xlab="Lag")
#matplot(xlag, resp_county_all, type="l", col=grey(0.5),lty=2,add=T)
abline(h=1)
lines(resp_cp_smk10,col=2,lwd=2)
legend ("top","Pooled (with 95%CI)",#,"First-stage region-specific"),
  lty=c(1,2),lwd=1.5,col=c(2,grey(0.7)),bty="n",inset=0.1,cex=0.8)
mtext(text=paste("10 ug/m^3 increase in GWR Smoke and All Respiratory Hospitalizations"))


# plot of cvd
plot.crosspred(cvd_cp_smk10, type="n", ylim=c(.8, 1.2), ylab="RR", xlab="Lag")
#matplot(xlag, resp_county_all, type="l", col=grey(0.5),lty=2,add=T)
abline(h=1)
lines(cvd_cp_smk10,col=2,lwd=2)
legend ("top","Pooled (with 95%CI)",#,"First-stage region-specific"),
  lty=c(1,2),lwd=1.5,col=c(2,grey(0.7)),bty="n",inset=0.1,cex=0.8)
mtext(text=paste("10 ug/m^3 increase in GWR Smoke and CVD Hospitalizations"))

```
