---
title: "Distributed Lag Model"
author: "Ryan_Gan"
date: "January 24, 2017"
output: html_document
---


## Purpose of Markdown File

Trying out distributed lag models using the 'dlnm' package. Loading libraries.

```{r libraries used, echo = F, message = F}

library(tidyverse) # data wrangle package
library(dlnm) # distributed lag models
library(lme4) # random-effects model
library(broom) # broom for tidy data from stats models
library(splines) # spline library
library(rmeta) # meta analysis
library(mvmeta) # multivariate meta analysis for pooling region-specific dlnm
```

Importing county-level time-series data.

```{r county time series import, echo = F, message = F}
# file path
file_path <- paste0("../../data/health_data/wa_2012_county_time_series.csv")

# read file
wash_ts_df <- read_csv(file_path)

summary(wash_ts_df)

# analysis dataframe
wash_jul_oct_ts_df <- wash_ts_df %>% 
  # restricting points
  filter(date >= "2012-07-01" & date <= "2012-10-31") %>% 
  # exclude 'Unknown' and 'not_wa_residence'
  filter(county != "not_wa_residence" & county != "Unknown") %>% 
  # create day of the week variable
  mutate(day = as.factor(weekdays(date))) %>% 
  # set missing outcome values to 0 since missing indicates no ER or urgent care 
  # visits on that date (which is reasonable in sparsely populated counties)
  mutate_each(funs(wo_miss = ifelse(is.na(.), 0, .)), n_obs:ra_n) %>% 
  # create a binary smoke indicator variable based smk variables >5, >10, and >15 units 
  mutate(wrf_smk5 = ifelse(wrf_smk_pm >= 5, 1, 0), 
         wrf_smk10 = ifelse(wrf_smk_pm >= 10, 1, 0),
         wrf_smk15 = ifelse(wrf_smk_pm >= 15, 1, 0), 
         geo_smk5 = ifelse(geo_smk_pm >= 5, 1, 0),
         geo_smk10 = ifelse(geo_smk_pm >= 10, 1, 0),
         geo_smk15 = ifelse(geo_smk_pm >= 15, 1, 0),
         krig_smk5 = ifelse(krig_smk_pm >= 5, 1, 0),
         krig_smk10 = ifelse(krig_smk_pm >= 10, 1, 0),
         krig_smk15 = ifelse(krig_smk_pm >= 15, 1, 0), 
         season = ifelse(date >= "2012-06-22" &  date <= "2012-09-22", "summer",
                  ifelse(date >= "2012-09-23" & date <= "2012-12-21", "fall",
                          "other"))) %>% 
  # rescale pm variables to 10 units
  mutate_each(funs(unit10 = ./10), wrf_pm:krig_smk_pm) %>% 
  select(1:2, 31:33, 36, 34:35, 38:44, 28:29, 30, 55, 58, 
         65, 64, 63) # subset and order variables used in analysis

```

## Visualizing Time-Series Data

First step is to visualize how both the health outcomes and our smoke concentrations vary over time.

### Health Outcomes for the Full Year of 2012

First plot contains a small-multiples plot of health outcomes of interest that aggregates all county-specific time-series for the entire state. There is no apparent pattern over the July 1st to October 31st timeseries for CVD outcomes. It's generally hard to observe these patterns in some of the sparsley populationed counties where most smoke occured, so I like to aggregate for the state to get an idea of the background counts.

```{r plot of events over time, echo = F, message = F, results="asis"}

# wrangle time-series dataframe to create small multiples of health outcomes
outcome_plot_df <- wash_ts_df %>% filter(county != "Unknown") %>% # remove unknown county
  select(1:8, 10:16) # select columns of health outcomes

wa_state_aggregate <- outcome_plot_df %>% select(-county) %>% 
  group_by(date) %>% 
  summarise_each(funs(sum(., na.rm = T))) %>% 
  gather(key = outcome, value = n, -date)

#summary(wa_state_aggregate)

ggplot(wa_state_aggregate, aes(x = date, y = n)) +
  geom_point() +
  facet_wrap(~outcome, scales = "free")
  
```

### Health Outcome Time Series from July 1st to October 31st 2012

Plot of the time trends in outcomes for just the study period of interest.

```{r plot of events jul to oct, echo = F, message = F, results="asis"}

# wrangle time-series dataframe to create small multiples of health outcomes
outcome_plot_df <- wash_ts_df %>% filter(county != "Unknown") %>% # remove unknown county
  select(1:8, 10:16) # select columns of health outcomes

wa_state_aggregate_jul_oct <- outcome_plot_df %>% select(-county) %>% 
  group_by(date) %>% 
  summarise_each(funs(sum(., na.rm = T))) %>% 
  gather(key = outcome, value = n, -date) %>% 
  filter(date >= "2012-07-01" & date <= "2012-10-31")

#summary(wa_state_aggregate)

ggplot(wa_state_aggregate_jul_oct, aes(x = date, y = n)) +
  geom_point() +
  facet_wrap(~outcome, scales = "free")

```  

It appears that there may be some temporal trends with some of the respiratory conditions. Asthma is a good example, where the number in the state is around 10 in July/August, but creeps up to as high as 30 in mid-September and early-October. Could part of that increase be due to wildfire smoke or is it part of a natural trend?


### County Smoke Concentrations Over Time

Most counties in the state are not impacted by smoke. However, some of the central counties do have the expected peak of high smoke concentrations around mid September.

```{r }
# check time series of smoke in each county
ggplot(wash_jul_oct_ts_df, aes(x = date, y = geo_smk_pm_unit10)) +
  geom_point() +
  facet_wrap(~county)
```


### Asthma and Smoke Time-Series Analysis

Counts of asthma claims tend to tick up a bit in the Fall, right around the state was affected by smoke. First step is to account for the temporal trends in asthma.

Plotting the asthma counts for the overall state and then each county for the July 1st to October 31st 2012 time series.

```{r study period asthma counts by county, message=F, echo=F, results="asis"}
# aggregated counts limited to July 1 to Oct 31st 2012
asthma_state_aggregate_jul_oct <- wa_state_aggregate %>% 
  filter(date >= "2012-07-01" & date <= "2012-10-31") %>% 
  filter(outcome == "asthma_n")

ggplot(asthma_state_aggregate_jul_oct, aes(x = date, y = n)) +
  geom_point() +
  ggtitle("washington asthma counts for the 2012 fire season")

# counties small multiple plot of asthma count
ggplot(wash_jul_oct_ts_df, aes(x = date, y = asthma_n_wo_miss)) +
  geom_point() +
  facet_wrap(~county) +
  ggtitle("asthma counts for the 2012 fire season by county")

```

It looks like in some of the bigger counties, asthma counts begin to increase in mid August. In King county (where Seattle is), the trend may even suggest that the peak is mid September (around the same time as the fire), and then beings to decrease. King county wasn't really affected by smoke, so it would be safe to say this is shape likely represents the background trend in asthma. It is hard or impossible to see trends in the more sparsely populated counties. Maybe this would be a good place to try and use empirical Bayes estimation? I should ask Ander.

There is evidence of overdispersion in each county. General Poisson regression may not be appropriate. 

Another issue is how to deal with any potential variation in asthma counts by county. Off the top of my head, it can either be ajusted for as a fixed effect, treated as a random effect, run each counties as seperate models and pool results in a meta-analysis. I'd also like to try an empirical Bayes approach.

### GLM Poisson without Natrual Spline

Exploring very basic models.

```{r basic glm regression}
# crude model with Poisson distribution
pois_mod <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10, wash_jul_oct_ts_df, 
            family="poisson")
# print mod
tidy(pois_mod)

# dispersion parameter for a poisson regression is assumed to be 1
summary(pois_mod)$dispersion

# crud model with Quasi-Poisson distribution for overdispersion
quasipois_mod <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10, wash_jul_oct_ts_df, 
            family="quasipoisson")
# print mod
tidy(quasipois_mod)
# dispersion parameter for quasipoisson
summary(quasipois_mod)$dispersion

# dispersion parameter for quasipoisson is >1, suggesting overdispersion

# beta estimate is not different, but standard error is higher in quasipoisson
```

Based on just these results, it's best to use the quasipoisson distribution to account for zero-inflated values. Beta estimate doesn't change much, but standard error is different enough where it could lead to different conclusions of just Poisson was used.

All subsequent models will use quasipoisson family. Next model adjusts for temperature and county

```{r glm adjusting for temperature}

# adjusted for temp
quasipois_mod <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10 + wrf_temp, 
                     wash_jul_oct_ts_df, family="quasipoisson")
# print mod
tidy(quasipois_mod)
# dispersion parameter
summary(quasipois_mod)$dispersion
# still overdispersed adjusting for temp

# adjusting for county as a fixed effect
quasipois_mod <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10 + wrf_temp + 
                          county, wash_jul_oct_ts_df, family="quasipoisson")
# print mod
tidy(quasipois_mod) 
# dispersion parameter
summary(quasipois_mod)$dispersion
# when accounting for county, dispersion is actually pretty close to 1
# suggests it may be okay to use a standard poisson distributions as
# long as county is accounted for

```

Summary from above models is that once county is accounted for, a 10 ug/m^3 increase in smoke PM2.5 is associated with a ~11% increase in asthma hospitalizations for ED or urgent care. Also, when county is accounted for, the dispersion parameter is approximaly 1. This suggests that it may be okay to use a Poisson distribution.

I'm going to try a random effect model with a Poisson distribution, as quasipoisson family does not exist for mixed model package.

```{r time series with a mixed model}

# asthma model where county is a random effect instead of fixed
mixed_mod <- glmer(asthma_n_wo_miss ~ geo_smk_pm_unit10 + wrf_temp + 
                  (1|county), wash_jul_oct_ts_df, family = "poisson")
# full output
summary(mixed_mod)

# reduced output
tidy(mixed_mod)

```

These results are almost identical to the model where county is treated as a fixed effect. Increases in a 10 ug/m^3 of wildfire smoke PM2.5 is associated with an ~11% increase in asthma hospitalizations.

### Pooled County-Specific Time Series 

This approch runs a time series for each county, then pools the results in a meta-analysis. I expect this will yield very similar results to the above results.


```{r meta of county time series}

county_mods <- wash_jul_oct_ts_df %>% 
  group_by(county) %>% # group by county
  do(geo_smk_beta = glm(asthma_n_wo_miss ~ geo_smk_pm_unit10 + wrf_temp, 
                        data = ., family="quasipoisson")) 
  

county_betas <- tidy(county_mods, geo_smk_beta) %>% 
  filter(term == "geo_smk_pm_unit10")

# meta analysis results for asthma and smoke
meta_results <- meta.summaries(county_betas$estimate, county_betas$std.error, 
                               method = "fixed")

summary(meta_results)
meta_results

```

This meta-analysis approach yields very similar results where there is an 11% increase in asthma hospitalizations associated with a 10 ug/m^3 increase in wildfire smoke PM2.5.

### GLM Poisson with Natrual Spline

I will now fit a natural spline using the 'splines' package to see if it improves the model. There are a couple ways to do it, but 

```{r asthma association with splines}
# aggregate outcomes for the entire state and average/median smoke estimates
outcomes_df <- wash_jul_oct_ts_df %>% 
  select(date, contains("wo_miss")) %>% 
  group_by(date) %>% 
  summarise_each(funs(sum(.)))

pm_df <- wash_jul_oct_ts_df %>% 
  select(date, wrf_temp, contains("pm")) %>% 
  group_by(date) %>% 
  summarise_each(funs(mean, median))

wash_outcome_counts_pm <- outcomes_df %>% 
  full_join(pm_df, by = "date") %>% 
  mutate(day = weekdays(date),
    season = ifelse(date >= "2012-06-22" &  date <= "2012-09-22", "summer",
      ifelse(date >= "2012-09-23" & date <= "2012-12-21", "fall", "other")))

summary(wash_outcome_counts_pm)

# define the function of the spline for the asthma trend for the entrie state
asthma_spl <- ns(wash_outcome_counts_pm$date, df = 3)
# general question. How does the spline know the trend is for asthma and not 
# other outcomes? very curious... I will need to look in to this

# model n asthma events with a spline with 7 df (knotches)
asthma_mod <- glm(asthma_n_wo_miss ~ asthma_spl, wash_outcome_counts_pm, 
              family="poisson") 

# check spline fit to asthma
summary(asthma_mod)

# join predicted number of events to a new df for plotting
asthma_pred_df <- wash_outcome_counts_pm %>% 
  select(date, asthma_n_wo_miss) %>% 
  mutate(expected_n = predict(asthma_mod, type="response"),
         res = residuals(asthma_mod, type="response"))

# plot predicted events over outcomes
ggplot(asthma_pred_df, aes(x = date, y = asthma_n_wo_miss)) +
  geom_point() +
  stat_smooth(aes(y = expected_n), se = F, method = "loess") +
  theme_bw()

# plot residuals
ggplot(asthma_pred_df, aes(x = date, y = res)) +
  geom_point() +
  geom_hline(yintercept = 0, colour = "blue") +
  theme_bw()


# run the regression models with the spline
# models for association between smoke pm and daily asthma count 
# unadjusted model
asthma_mod_unadj <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10_median + 
                        wrf_temp_mean + day + season, 
                        wash_outcome_counts_pm, family="quasipoisson")
summary(asthma_mod_unadj)

# controlling for seasonality with spline
asthma_mod_adj <- update(asthma_mod_unadj,.~.+ asthma_spl)
summary(asthma_mod_adj)

```

When using a spline, there is no association between increasing smoke exposure and asthma counts. However, I had to aggregate to the state.

## Distributed Lag Models

General interpretation would be for a 1 ug/m^3 increase in smoke PM2.5, asthma events increase by 1%.

```{r simple dlm}

cb1.pm <- crossbasis(wash_ts_df$geo_wt_pm, lag = 15, argvars=list(fun="lin"),
                     arglag=list(fun="poly", degree=3))

summary(cb1.pm)

model1 <- glm(asthma_n ~ cb1.pm ,
              family=quasipoisson(), wash_ts_df)

summary(model1)


pred1.pm <- crosspred(cb1.pm, model1, at=0:20, bylag=0.2, cumul=TRUE)

plot(pred1.pm, "slices", var=10, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
main="Association with a 1-unit increase in PM2.5")


plot(pred1.pm, "slices", var=10, col=2, cumul=TRUE, ylab="Cumulative RR",
main="Cumulative association with a 1-unit increase in PM2.5")



```

May not be appropriate to aggregate for the entrie state. Just based on outcome and smoke plots, it's clear that there is heterogenity by county.

Gasparrini and Armstrong published an extension of their original distributed lag model work that is a two-stage approach that can pool the results from each region. Perfect for this study.

```{r chelan dlm, eval = F}
# make df
chelan_df <- wash_jul_oct_ts_df %>% 
  filter(county == "Chelan")

summary(chelan_df)

# define arguments and lists for cross-basis definition

# sparse counts of asthma visits. i'll try dlm on all resp and cvd outcomes

cb.smk <- crossbasis(chelan_df$geo_smk_pm_unit10, # vector of all smk pm vals in chelan 
                     lag = 10, # max lag days
                     argvar=list(fun="lin"), # assuming linear relationship 
                     arglag=list(fun="poly", degree=3)) # assuming polynomial lag function

?crossbasis
# wrf temp
cb1.temp <- crossbasis(chelan_df$wrf_temp, lag = 3, 
                     argvar=list(df=5), arglag=list(fun="strata", breaks=1))

summary(cb1.pm)
summary(cb1.temp)

# standard model
summary(glm(asthma_n_wo_miss ~ geo_smk_pm_unit10 + wrf_temp, 
            family = quasipoisson(), chelan_df))

# distributed lag model
model1 <- glm(asthma_n_wo_miss ~ cb1.pm + cb1.temp + day, family=quasipoisson(), chelan_df)

summary(model1)


pred1.pm <- crosspred(cb1.pm, model1, at=0:20, bylag=0.2, cumul=TRUE)

plot(pred1.pm, "slices", var=10, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
main="Association with a 10-unit increase in PM2.5")


plot(pred1.pm, "slices", var=10, col=2, cumul=TRUE, ylab="Cumulative RR",
main="Cumulative association with a 10-unit increase in PM2.5")

# random effects model for the whole state
# sparse counts of asthma visits. i'll try dlm resp and cvd
cb1.pm <- crossbasis(wash_jul_oct_ts_df$geo_smk_pm_unit10, lag = 15, 
                     argvar=list(fun="lin"), arglag=list(fun="poly", degree=5))

# wrf temp
cb1.temp <- crossbasis(wash_jul_oct_ts_df$wrf_temp, lag = 3, 
                     argvar=list(df=5), arglag=list(fun="strata", breaks=1))

summary(cb1.pm)
summary(cb1.temp)

# distributed lag model
model1 <- glm(asthma_n_wo_miss ~ cb1.pm + cb1.temp + county, family=quasipoisson(), 
              wash_jul_oct_ts_df)

summary(model1)

# plots
pred1.pm <- crosspred(cb1.pm, model1, at=0:20, bylag=0.2, cumul=TRUE)

plot(pred1.pm, "slices", var=10, col=3, ylab="RR", ci.arg=list(density=15,lwd=2),
main="Association with a 10-unit increase in PM2.5")


plot(pred1.pm, "slices", var=10, col=2, cumul=TRUE, ylab="Cumulative RR",
main="Cumulative association with a 10-unit increase in PM2.5")

```


```{r dlm in a mixed model, eval = F}
# random effects model for the whole state
# sparse counts of asthma visits. i'll try dlm resp and cvd
cb_smk <- crossbasis(wash_jul_oct_ts_df$geo_smk_pm_unit10, lag = 10, 
                     argvar=list(fun="lin"), arglag=list(fun="poly", degree=5))

# wrf temp
cb_temp <- crossbasis(wash_jul_oct_ts_df$wrf_temp, lag = 3, 
                      argvar=list(df=5), arglag=list(fun="strata", breaks=1))

summary(cb_smk)
summary(cb_temp)

# distributed lag model
mixed_mod <- glmer(asthma_n_wo_miss ~ cb_smk + cb_temp + 
                  (1|county), wash_jul_oct_ts_df, family = "poisson")
# full output
summary(mixed_mod)
```
