---
title: "Select ZIP PM2.5 Timeseries"
author: "Ryan_Gan"
date: "June 5, 2017"
output: html_document
---

### RE: Matthew Cadlec inquiry on smoke PM~2.5~ values 

This document contains our geo-weighted ridge regression (GWR) summary statistics and time series plots for select ZIP codes from cities around Washington state during the 2012 wildfire season.

Question from Matthew Cadlec:

> *I’m curious what the ranges and means of daily PM2.5 concentrations were in the cities impacted by wildfire smoke as calculated by your geographically weighted ridge regressions.*  

```{r setup, message=F, warning=F}
# load library
library(tidyverse)

# read in pm values for smoke
smoke_path <- paste0("../../data/pm_data/")

zip_pm <- read_csv(paste0(smoke_path, "zip_pm_to_merge_with_chars.csv")) %>% 
  filter(complete.cases(.))
```

### Subset of ZIP codes

I'm subsetting our larger ZIP code population-weighted smoke values to the following ZIP codes as representatives of the cities in which they were drawn from.

- Seattle: 98104
- Tacoma: 98402
- Wenatchee: 98802
- Yakima: 98902
- Kennewick: 99336
- Spokane: 99201

```{r subset zips, message=F, warning=F, results="asis"}
zip_sub <- zip_pm %>% 
  filter(ZIPCODE %in% c(98104, 98402, 98802, 98902, 99336, 99201)) %>% 
  # assign city names to zips
  mutate(city = ifelse(ZIPCODE == 98104, "Seattle",
                ifelse(ZIPCODE == 98402, "Tacoma",
                ifelse(ZIPCODE == 98802, "Wenatchee",
                ifelse(ZIPCODE == 98902, "Yakima",
                ifelse(ZIPCODE == 99336, "Kennewick", "Spokane")))))) %>% 
  # subset variables to GWR 
  select(city, ZIPCODE, date, geo_wt_pm, background_pm, geo_smk_pm)

# summarize values requested
city_summary <- zip_sub %>% 
  group_by(city) %>% 
  summarise_each(funs(mean = mean, median= median, 
                      min = min, max = max), geo_smk_pm) 

# kable table
knitr::kable(city_summary, caption = paste0("Summary statistics (mean, median,",
"min, max) over the timeseries for select cities in Washington"))

```

### Timeseries plot for each city (ZIP code)

I've include 3 methods of estimating PM2.5. Background method represents ambient PM~2.5~ levels without smoke present. Geo Wt PM is the untransformed estimate of PM~2.5~ via the GWR method. Geo smk pm subtracts the background level off. 

```{r timeseries, message=F, warning=F, results="asis"}
# small multiples dataframe
city_ts_df <- zip_sub %>% 
  select(-ZIPCODE) %>% 
  gather(key = pm_method, value = pm, -city, -date)

  
# small multiples plot
ggplot(city_ts_df, aes(x=date, y=pm, colour = pm_method)) +
  geom_line() +
  facet_wrap(~city) +
  ylab("PM2.5 ug/m^3") +
  xlab("Date") + 
  theme_bw()

```