---
title: 'Washington: County-Level Wildfire Smoke and Health Outcomes'
author: "Ryan Gan"
date: "October 3, 2016"
output: html_document
---

## Overview

This document contains county-level results for the association between wildfire smoke PM~2.5~ and cardiovascular (CVD) and respiratory health outcomes. These results will be paired with Rish's county-level mortality analyses for a third manuscript. Herein, I compare different PM~2.5~ estimation methods described in further detail below, as well as contrasting results where county of residence is used to assign population-weighted PM~2.5~ estimates to results where zip code is used to assign population-weighted PM~2.5~. The time period of interest is July 1st through October 31st 2012, meaning I consider all health outcome events during this time period for all study designs. Rish has a slightly narrower period from the 15th of August to the 15th of October.

```{r setup, include=FALSE}
# loading libraries used
library(tidyverse) # import tidyverse
library(survival) # for conditional logistic regression
library(htmlTable) # table
library(rmeta) # for meta analysis
library(lme4) # random-effects model
library(broom) # broom for tidy data from stats models
```

```{r casecross data import, include = F, echo = FALSE}
# Set working directory and read in files --------------------------------------
# Direct path 
path <- paste0("C:/Users/RGan/Documents/git_local_repos/wildfire/",
               "wildfire_washington/analysis/analysis_data")

# Infile case-crossover dataframes that contain both zip and county pop-wt pm2.5
# resp exacerbations
resp_casecross <- read_csv(paste(path, "resp1_jul_to_oct_casecross.csv", 
                                 sep = "/"))
# asthma
asthma_casecross <- read_csv(paste(path, "asthma1_jul_to_oct_casecross.csv", 
                                   sep = "/"))
# copd 
copd_casecross <- read_csv(paste(path, "copd1_jul_to_oct_casecross.csv", 
                                 sep = "/"))
# copd exacerbations
copd_ex_casecross <- read_csv(paste(path, "copd_ex1_jul_to_oct_casecross.csv",
                                    sep="/"))
# pneum or bronchitis
pneum_casecross <- read_csv(paste(path, "pneum1_jul_to_oct_casecross.csv",
                                  sep="/"))
# acute bronchitis
acute_bronch_casecross <- read_csv(paste(path, 
                                   "acute_bronch1_jul_to_oct_casecross.csv",
                                         sep = "/"))
# cvd
cvd_casecross <- read_csv(paste(path, "cvd1_jul_to_oct_casecross.csv",
                                sep="/"))
# arrhythmia
arrhythmia_casecross <- read_csv(paste(path, "arrhythmia1_jul_to_oct_casecross.csv",
                                 sep="/"))
# cerebral vascular
cereb_vas_casecross <- read_csv(paste(path, "cereb_vas1_jul_to_oct_casecross.csv",
                                sep="/"))
# heart failure
hf_casecross <- read_csv(paste(path, "hf1_jul_to_oct_casecross.csv", 
                               sep="/"))
# ischemic heart disease
ihd_casecross <- read_csv(paste(path, "ihd1_jul_to_oct_casecross.csv",
                                sep="/"))
# myo infarc
mi_casecross <- read_csv(paste(path, "mi1_jul_to_oct_casecross.csv", sep="/"))
# RA
ra_casecross <- read_csv(paste(path, "ra1_jul_to_oct_casecross.csv", sep="/"))
# broken arm
broken_arm_casecross <- read_csv(paste(path, "broken_arm1_jul_to_oct_casecross.csv",
                                       sep="/"))

# time-serieis dataframe
wash_ts_df <- read_csv(paste(path, "wa_2012_county_time_series.csv", sep="/"))

```

## Time-Series Study (County)

County-level time-series' use daily counts of the outcome of interest and daily estimates of the exposure (PM~2.5~), where the day is the unit is observation. An assumption is that the underlying population doesn't vary much day to day, therefore you only need a numerator and don't need a denominator. Note, I didn't try zip code-level time-series as I think the spatial area is too small.


### Trends in Outcomes Over the Year

First step is to see if there are seasonal trends in the outcome within each county over the year. It is important to look out for any sort of clear pattern of the outcome over the year. Best example is the seasonal variation in asthma counts. In the interest of being concise, I've only printed asthma here, although I've looked at each outcome (first chunk of R code is turned off via eval = False), and really the only outcome that exhibits any type of seasonal variation is asthma. Furthermore, only King county exhibits any sort of noticeable trend. Thus, I don't think it's necessary to fit any types of splines to account for seasonal trends for the time-series models.

```{r outcome trends, eval = F, echo = F, results='asis'}

# create a dataframe, filling in missing values for the whole year
wash_ts_no_miss_df <- wash_ts_df %>% 
  # exclude 'Unknown' and 'not_wa_residence'
  filter(county != "not_wa_residence" & county != "Unknown") %>% 
  # set missing outcome values to 0 since missing indicates no ER or urgent care 
  # visits on that date (which is reasonable in sparsely populated counties)
  mutate_each(funs(wo_miss = ifelse(is.na(.), 0, .)), n_obs:ra_n)

#summary(wash_ts_no_miss_df)

outcome_list <- c('All ER or Urgent', 'All Respiratory', 'Asthma', 'COPD', 
                  'COPD Exacerbation', 'Pneumonia', 'Acute Bronchitis', 
                  'Cardiovascular Disease', 'Arrhythmia', 'Cerebrovascular Disease', 
                  'Heart Failure', 'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Rheumatoid Arthritis', 'Broken Arm')

which(colnames(wash_ts_no_miss_df)=='n_obs_wo_miss')
# plot loop
for(i in 30:44){

var_name <- colnames(wash_ts_no_miss_df[i])
outcome_name <- outcome_list[i-29]


plot <- ggplot(wash_ts_no_miss_df, aes_string(x= 'date', y = var_name)) +
  geom_jitter() +
  facet_wrap(~county, scales = "free", ncol = 4) +
  ggtitle(paste('Number of ', outcome_name, ' Events by County in 2012',  sep = '')) +
  ylab('Count of Events') +
  xlab('Date') +
  scale_x_date(date_labels = "%b") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.title.y = element_text(angle = 90))  


print(plot)

  
}

```

```{r asthma trends, echo = F, results='asis'}

# create a dataframe, filling in missing values for the whole year
wash_ts_no_miss_df <- wash_ts_df %>% 
  # exclude 'Unknown' and 'not_wa_residence'
  filter(county != "not_wa_residence" & county != "Unknown") %>% 
  # set missing outcome values to 0 since missing indicates no ER or urgent care 
  # visits on that date (which is reasonable in sparsely populated counties)
  mutate_each(funs(wo_miss = ifelse(is.na(.), 0, .)), n_obs:ra_n)

#summary(wash_ts_no_miss_df)


plot <- ggplot(wash_ts_no_miss_df, aes_string(x= 'date', y = 'asthma_n_wo_miss')) +
  geom_jitter() +
  facet_wrap(~county, scales = "free") +
  ggtitle('Number of Asthma Events by County in 2012') +
  ylab('Count of Asthma') +
  xlab('Date') +
  scale_x_date(date_labels = "%b") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.title.y = element_text(angle = 90))  

print(plot)

```


The analyses I used were a generalized mixed model, where the dependent variable was the daily outcome counts, regressed on continuous PM~2.5~, assuming a linear relationship (this assumption will be explored in later analyses). The models are adjusted for daily temperature (Kelvin), and day of the week, which I found didn't add much to base models, but I adjusted for as this is a common adjustment in other epi studies. I also treated county as a random intercept, allowing it to vary county to county. Note that the random effects model produces almost identical regression estimates for smoke as if county was adjusted for as a fixed effect.

The results can be interpreted as you would a risk ratio or odds ratio from the time-stratified case-cross over design.


```{r time series rando effect, echo = F, warning= F, results='asis'} 

# analysis dataframe
wash_jul_oct_df <- wash_ts_df %>% 
  # restricting points
  filter(date >= "2012-07-01" & date <= "2012-10-31") %>% 
  # exclude 'Unknown' and 'not_wa_residence'
  filter(county != "not_wa_residence" & county != "Unknown") %>% 
  # create day of the week variable
  mutate(day = as.factor(weekdays(date))) %>% 
  # set missing outcome values to 0 since missing indicates no ER or urgent care 
  # visits on that date (which is reasonable in sparsely populated counties)
  mutate_each(funs(wo_miss = ifelse(is.na(.), 0, .)), n_obs:ra_n) %>% 
  # rescale pm variables to 10 units
  mutate_each(funs(unit10 = ./10), wrf_pm:krig_smk_pm) %>% 
  select(1:2, 30:45, 28:29, 48, 53:55) # subset to variables used in analysis

# finding columns to retain
#which(colnames(wash_jul_oct_df)=='wrf_smk_pm_unit10')

#glimpse(wash_jul_oct_df)
# outcomes to loop through
outcome_list <- c('All ER or Urgent', 'All Respiratory', 'Asthma', 'COPD', 
                  'COPD Exacerbation', 'Pneumonia', 'Acute Bronchitis', 
                  'Cardiovascular Disease', 'Arrhythmia', 'Cerebrovascular Disease', 
                  'Heart Failure', 'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Rheumatoid Arthritis', 'Broken Arm')

# find list of county names
county_names <- c("Adams", "Asotin", "Benton", "Chelan", "Clallam", "Clark", 
                  "Columbia", "Cowlitz", "Douglas", "Ferry", "Franklin", 
                  "Garfield", "Grant", "Grays Harbor", "Island", "Jefferson", "King",
                  "Kitsap", "Kittitas", "Klickitat", "Lewis", "Lincoln", "Mason", 
                  "Okanogan", "Pacific", "Pend Oreille", "Pierce", "San Juan", 
                  "Skagit", "Skamania", "Snohomish", "Spokane", "Stevens", 
                  "Thurston", "Wahkiakum", "Walla Walla", "Whatcom", "Whitman", 
                  "Yakima")


# method names to assign
method_name <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke',  'Global Smoke', 
                 'Kriging Smoke')

# empty matrix for table
point_estimates <- matrix(nrow = 4, ncol = 4, byrow = T)
  
colnames(point_estimates) <- c('pm_method', 'risk_ratio', 
                               'lower95', 'upper95')

# find column number
# which(colnames(wash_jul_oct_df)=='ra_n_wo_miss')  

# loop through outcomes
for(i in 4:18){

  # variable to model 
  outcome_name <- outcome_list[i-3]

  # loop through pm estimation methods
  for(j in 21:24){

  var_name <- method_name[j-20]
  
  # random effects model where county is treated as a random intercept 
  mixed_mod_adj <- glmer(wash_jul_oct_df[[i]] ~ wash_jul_oct_df[[j]] + day +
                         wrf_temp + (1|county), wash_jul_oct_df, family = "poisson")
  
  model_df <-tidy(mixed_mod_adj)

  # populate matrix
  row_n <- j-20
  # method
  point_estimates[row_n, 1] <- var_name
  # rate ratio
  point_estimates[row_n, 2] <- round(exp(model_df[2,2]), 3)
  # 95% lower bound
  point_estimates[row_n, 3] <- round(exp((model_df[2,2]) -
                                    1.96*(model_df[2,3])), 3)
  # 95% upper bound
  point_estimates[row_n, 4] <- round(exp((model_df[2,2]) +
                                    1.96*(model_df[2,3])), 3)

  }

  # save point estimates as a dataframe
  point_est_df <- data.frame(point_estimates)

    
tab <- htmlTable(txtRound(point_est_df, digits = 3, 2:4), 
        caption = paste("Time-Series: Association between a 10 ug/m^3 in PM2.5 (county population-weighted) and", outcome_name, sep = " "),
        header = c("Method", "Risk Ratio&dagger;", "Lower", "Upper"),
        cgroup = c("", "95% CI"), # spanner
        n.cgroup = c(2, 2),
        padding.rgroup = "&nbsp;&nbsp;",
        css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
        align = "lccc", # column alignment,
        tfoot="&dagger; Adjusted for temperature and day of the week, accounting for county as a random intercept."
            ) # end table
  
print(tab)
  
# model_adj <- glm(asthma_n_wo_miss ~ geo_smk_pm_unit10 + wrf_temp + as.factor(county), wash_jul_oct_df, family=quasipoisson)
# summary(model_adj)
}


```

### Time-Series Results (County)

The following tables contain the risk ratio estimates for the association between a linear increase in wildfire smoke PM~2.5~ and the health outcomes of interest. Using asthma as our example, we'd conclude that between July and October in 2012, a 10 unit increase wildfire smoke PM~2.5~, the risk for asthma increased by ~12% to ~15%, depending on the smoke estimate used. All smoke estimation methods were statistically significant for asthma. 

## Time-Stratified Case-Crossover (County)

### Symmetric bi-directional multiple observations, including only 3rd referent observation before and after the admission. PM~2.5~ assignment based on county of residence. I only use subject with single observations to avoid problems that arrise when subjects have multiple observations.

In the interest of producing a concise document, I'm only presenting using the 3rd referent observation as this seems to be a good compromise accounting for 'recovery period' and staying within a reasonable amount of time from the observed event for referent periods.  

```{r county symmetrical and bi-directional, echo = F, results='asis'} 
# note above, results = 'asis' needed to print htmlTables
# Analyses ---------------------------------------------------------------------
# look up admit type, may want to subset to specifc admit
# ADM_TYPE variable: 1 = emergency, 2 = urgent, 3 = elective, 4 = newborn
# 5 = trauma, 9 = info not available

# dataframe list
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, copd_ex_casecross, 
  pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
  cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
  ra_casecross, broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'COPD Exacerbation',
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Rheumatoid Arthritis', 'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 
                 'Global Smoke', 'Kriging Smoke')


# Producing conditional logit model estimates loop -----------------------------
for(i in 1:length(df_list)){

  df_to_loop <- data.frame(df_list[i])
  outcome <- colnames(df_to_loop[3])

  outcome_name <- outcome_list[i]
  #glimpse(df_to_loop)

  # find the column numbers for the county-level temp and pbl
  # which(colnames(df_to_loop)=="day") 
  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:16, 94:95, 151:155)]
  # extract pm values and divide by 10 and ordered
  # which(colnames(df_to_loop)=="krig_smk_pm_county") # code to find column numbers
  pm_estimates_df <- df_to_loop[, c(86, 91:93)]/10  # create 10 unit increases
  # bind columns back together
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    filter(!is.na(wrf_smk_pm_county)) %>% 
    # limit to emergency or urgent care
    filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (date - date_admit)/7) %>% 
    group_by(PATIENTID) %>% 
    # identifies the min and max observations, and sets it to the minimum n of obs
    mutate(min_obs_diff = abs(min(obs_diff_admission)),
           max_obs_diff = abs(max(obs_diff_admission)),
           obs_limit = min(c(min_obs_diff, max_obs_diff)),
           keep_obs = ifelse(abs(obs_diff_admission) <= obs_limit, 1, 0),
           # excludes the 1st counterfactual observation bf/after admission
           keep_obs2 = ifelse(obs_diff_admission == 0 | obs_diff_admission == 3 |
                              obs_diff_admission == -3, 1, 0),
           # get rid of people with single observations
           keep_obs3 = ifelse(obs_limit >= 3, 1, 0)) %>% 
    filter(keep_obs == 1 & keep_obs2 == 1 & keep_obs3 == 1)

  # second loop to run a model for each pm estimation method
  
  # empty df for table
  table_df <- data.frame()
  
  # empty matrix
  point_estimates <- matrix(nrow = 4, ncol = 8, byrow = T)
  
  colnames(point_estimates) <- c('pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')
  #glimpse(df_analysis)
  
    for(j in 24:27){

      # variable to model 
      var_name <- colnames(df_analysis[j])

      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_county + 
                    strata(PATIENTID), df_analysis)
      # might want to consider 'broom' package in the future to make it easier to get 
      # model estimates.

      # populate matrix
      row_n <- j-23

      point_estimates[row_n, 1] <- var_name
      point_estimates[row_n, 2] <- mod$n
      point_estimates[row_n, 3] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 4] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 5] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 7] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,5], 4)
  
      # save point estimates as a dataframe
      point_est_df <- data.frame(point_estimates)
      assign(paste(outcome, "point_est", sep = '_'), point_est_df) 
    }
  
  table_df <- cbind(data.frame(method_list), point_est_df[, 2:6])
  
  tab <- htmlTable(txtRound(table_df, digits = 3, 1:3 ), 
           caption = paste("Association between a 10 ug/m^3 in PM2.5 (county population-weighted) and",
                           outcome_name, sep = " "),
           header = c("Method", "Obs.", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           cgroup = c("", "95% CI"), # spanner
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "lccccc", # column alignment,
           tfoot="&dagger; Adjusted for temperature, accounting for subject. Symmetric bi-direction; including only the 3rd counterfactual observations."
            ) # end table
  
  print(tab)
}

```

### Time-Series Results (Zipcode)

I'm including the time-stratified case-crossover results using a referent period of 3 observations (weeks) before or after the observed event, where zip code as the spatial unit to assign PM~2.5~. This would offer a direct comparison to county-level estimates.

```{r zip symmetrical and bi-directional, echo = F, results='asis'} 
# note above, results = 'asis' needed to print htmlTables
# Analyses ---------------------------------------------------------------------
# look up admit type, may want to subset to specifc admit
# ADM_TYPE variable: 1 = emergency, 2 = urgent, 3 = elective, 4 = newborn
# 5 = trauma, 9 = info not available

# dataframe list
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, copd_ex_casecross, 
  pneum_casecross, acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
  cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
  ra_casecross, broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'COPD Exacerbation',
                  'Pneumonia', 'Acute Bronchitis', 'Cardiovascular Disease',
                  'Arrhythmia', 'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction',
                  'Rheumatoid Arthritis', 'Broken Arm')

method_list <- c('WRF-Chem Smoke', 'Geo-Weighted Smoke', 
                 'Global Smoke', 'Kriging Smoke')


# Producing conditional logit model estimates loop -----------------------------
for(i in 1:length(df_list)){

  df_to_loop <- data.frame(df_list[i])
  outcome <- colnames(df_to_loop[3])

  outcome_name <- outcome_list[i]
  #glimpse(df_to_loop)

  # extract covariates from dataframe
  covariates_df <- df_to_loop[, c(1:16, 27:28, 151:155)]
  # extract pm values and divide by 10 and ordered
  #which(colnames(df_to_loop)=="wrf_smk_pm_zip") # code to find column numbers
  pm_estimates_df <- df_to_loop[, c(19, 24:26)]/10  # create 10 unit increases
  # bind columns back together
  df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
    filter(!is.na(wrf_smk_pm_zip)) %>% 
    # limit to emergency or urgent care
    filter(ADM_TYPE == 1 | ADM_TYPE == 2) %>%
    # the following code makes sure that the counterfactual values retained are 
    # symetric in that number of obs before = number of obs after
    mutate(obs_diff_admission = (date - date_admit)/7) %>% 
    group_by(PATIENTID) %>% 
    # identifies the min and max observations, and sets it to the minimum n of obs
    mutate(min_obs_diff = abs(min(obs_diff_admission)),
           max_obs_diff = abs(max(obs_diff_admission)),
           obs_limit = min(c(min_obs_diff, max_obs_diff)),
           keep_obs = ifelse(abs(obs_diff_admission) <= obs_limit, 1, 0),
           # excludes the 1st counterfactual observation bf/after admission
           keep_obs2 = ifelse(obs_diff_admission == 0 | obs_diff_admission == 3 |
                              obs_diff_admission == -3, 1, 0),
           # get rid of people with single observations
           keep_obs3 = ifelse(obs_limit >= 3, 1, 0)) %>% 
    filter(keep_obs == 1 & keep_obs2 == 1 & keep_obs3 == 1)

  # second loop to run a model for each pm estimation method
  
  # empty df for table
  table_df <- data.frame()
  
  # empty matrix
  point_estimates <- matrix(nrow = 4, ncol = 8, byrow = T)
  
  colnames(point_estimates) <- c('pm_method', 'n', 'n_events', 'odds_ratio', 
                                 'lower95', 'upper95', 'se', 'p_val')
  #glimpse(df_analysis)
  
    for(j in 24:27){

      # variable to model 
      var_name <- colnames(df_analysis[j])

      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + 
                    strata(PATIENTID), df_analysis)
      # might want to consider 'broom' package in the future to make it easier to get 
      # model estimates.

      # populate matrix
      row_n <- j-23

      point_estimates[row_n, 1] <- var_name
      point_estimates[row_n, 2] <- mod$n
      point_estimates[row_n, 3] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 4] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 5] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 6] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 7] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 8] <- round(summary(mod)$coefficient[1,5], 4)
  
      # save point estimates as a dataframe
      point_est_df <- data.frame(point_estimates)
      assign(paste(outcome, "point_est", sep = '_'), point_est_df) 
    }
  
  table_df <- cbind(data.frame(method_list), point_est_df[, 2:6])
  
  tab <- htmlTable(txtRound(table_df, digits = 3, 1:3 ), 
           caption = paste("Association between a 10 ug/m^3 in PM2.5 (zip population-weighted) and",
                           outcome_name, sep = " "),
           header = c("Method", "Obs.", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           cgroup = c("", "95% CI"), # spanner
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "lccccc", # column alignment,
           tfoot="&dagger; Adjusted for temperature, accounting for subject. Symmetric bi-direction; including only the 3rd counterfactual observations."
            ) # end table
  
  print(tab)
}

```

### Time-Stratified Case-Crossover Results (Zipcode)

In general, results are very similar to county assignment of PM~2.5~, likely as the population-weighted averaging still emphasizes the PM~2.5~ values where most people live. This is reassuring, and has broader implications for use of county-level data, namely if you can accout and assign environmental estimates based on where most people live, this can reduce exposure misclassification based on spatial unit of assignment. Whether the individual subject is actually exposed, such as if they spend most time indoors, or work in another county, is exposure misclassification that cannot be addressed in this study.

## Summary 

Time-series, county-level case-crossover, and zip-code level case-crossover all lead to similar conclusions regarding asthma and COPD. When the time-series design converges with the case-crossover design, we can be pretty confident in our results. I also think it's okay to add a couple referent periods (3 and 4 is ideal I think), but we can go with only using the 3rd referent period for now.

## To Do: 
1. Implement Rish's 1/2 month referent period case-crossover design after clarifying exactly how to set it up.
2. Run model with only smoke estimates in the model.
3. Try smoke interaction term.

## Future Work:
1. Defining appropriate referent period for each health outcome.
2. Evaluate non-linear trends between PM~2.5~ and health outcomes.

